Debiased Calibration for Generalized Two-Phase Sampling

### 2024-05-06 final/reports/prelim_paper.tex
Added Topic 3 methods.

The method of \cite{kwon2024debiased} requires known finite population 
calibration levels. It does not handle the
two-phase setup where we need to estimate the finite population total of $\bf x$
from the Phase 1 sample. In the rest of the section, we extend this method to 
two phase sampling so that we have a valid 
estimator when including estimated Phase 1 weights with appropriate variance
estimation.

@@

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (1, x_{1i}, x_{2i})^T$,
    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}

@@

Now we consider the sampling mechanism known as non-nested two phase sampling 
(\cite{hidiroglou2001double}). In the last section, we considered two phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
is a separate independent sample of the same population. Like traditional two
phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.

Whereas the classical two phase estimator uses a single Horvitz-Thompson
estimator of the Phase 1 sample to construct estimates for calibration totals,
in the non-nested two phase sample we have two independent Horvitz-Thompson
estimators of the total of $\bf X$,

@@

where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two phase
sample this regression estimator can be viewed as the solution to the following
calibration equation

@@

We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
non-nested two phase sampling case where we use a combined estimate 
$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
the finite population.

@@

The methodology for the non-nested two phase sample is very similar to the setup
described as part of Topic 1. Given a strictly convex differentiable function,
$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve

@@

We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
setup we minimize the Lagrangian,

@@

\section{Topic 3: Multi-source Two-Phase Sampling}

>>>>

The method of \cite{kwon2024debiased} requires known finite population 
calibration levels. It does not handle the
two-phase setup where we need to estimate the finite population total of $\bf x$
from the Phase 1 sample. In the rest of the section, we extend this method to 
two-phase sampling so that we have a valid 
estimator when including estimated Phase 1 weights with appropriate variance
estimation.

@@

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two-Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (1, x_{1i}, x_{2i})^T$,
    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}

@@

Now we consider the sampling mechanism known as non-nested two-phase sampling 
(\cite{hidiroglou2001double}). In the last section, we considered two-phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
is a separate independent sample of the same population. Like traditional two
phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
observations of $(\bf X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
of observations of $(\bf X_i, Y_i)_{i = 1}^{n_2}$.

Whereas the classical two-phase estimator uses a single Horvitz-Thompson
estimator of the Phase 1 sample to construct estimates for calibration totals,
in the non-nested two-phase sample we have two independent Horvitz-Thompson
estimators of the total of $\bf X$,

@@

where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the two-phase
sample this regression estimator can be viewed as the solution to the following
calibration equation

@@

We extend the debiased calibration estimator of \cite{kwon2024debiased} to the
non-nested two-phase sampling case where we use a combined estimate 
$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
the finite population.

@@

The methodology for the non-nested two-phase sample is very similar to the setup
described as part of Topic 1. Given a strictly convex differentiable function,
$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve

@@

We can constuct a non-nested two-phase estimator $\hat Y_{NNE}$ for $Y_N$ where
$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
Equation~\ref{eq:nnopt}. Like the classical two-phase approach, to solve this
setup we minimize the Lagrangian,

@@

\section{Topic 3: Multi-Source Two-Phase Sampling}

When considering non-nested two-phase sampling, we focused on the case of having
two samples. Now, we consider incorporating more than two independent samples
together with a debiasing constraint. First, we consider the case in which we
want to estimate the $\theta = E[Y]$ and $Y$ is only observed in one sample.
Then we will consider the case where $Y$ may be in multiple samples.

Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
samples with elements $\bf X^{(0)}$ observed in $A_0$ and $\bf X^{(m)}$ observed
in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we assume that
each survey is sampling independently from the same population sampling frame.

The traditional multi-source approach (\cite{kim2024statistics}) is to use
generalized least squares (GLS) to obtain an optimal estimator of $X$ from the
samples $A_1, \dots, A_M$. Then we can incorporate this information in the
estimation of $\theta$ by using the following estimate

$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$

\begin{equation}\label{eq:mstps}
\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } \sum_{i \in
A_0} w_i \bf x_i = \hat X_{GLS}^{(0)}
\end{equation}

where $\hat X_{GLS}^{(0)}$ is the GLS estimate of $X$ for all of the
$X$-variables measured in sample $A_0$, and $G$ is a generalized entropy
function.

% Outline
% Instead of incorporating data from one additional source, we want to combine
% data from multiple sources
% * Use GLS (generalized from effective sample size)
% * We need to add debiasing constraint 
%   * What do we do if we have multiple surveys for which we observe Y?
%   * How do we incorporate the selection probability?

In order to have a design consistent estimator that incorporates information from
samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraints,

$$ \sum_{i \in A_0} w_{i} \bf x_i q_i = \sum_{i \in U} g(d_i) q_i $$

where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
sample $A_0$. This yields the optimization problem:

\begin{equation}\label{eq:dc3}
  \hat w = \argmin_w \sum_{i \in A_0} G(w_i) q_i + \lambda\left(\hat T - \sum_{i
  \in A_0} w_i \bf z_i\right)
\end{equation}

where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
in which multiple samples can observe $Y$. In this case, assume that we have
samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
observed in samples $A_m$ and $B_m$ respectively. Then the final
estimator\footnote{This is sort of like stratified sampling with unknown strata.}
is

$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$

and $\hat w_{im}$ solves

\begin{equation}\label{eq:dc32}
  \hat w = \argmin_w \sum_{i \in A_m} G(w_i) q_i + \lambda\left(\hat T_m - \sum_{i
  \in A_m} w_i \bf z_i\right)
\end{equation}

where $\hat T = (\hat{\bf X}^{(m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
are the design weights for $A_m$.

\subsection{Simulation Study 3.1}

\subsection{Simulation Study 3.2}
